# Network Analysis

```{r setup}
library(STRINGdb)
library(graph)
library(RCytoscape)
library(org.Hs.eg.db)
library(hgu133plus2.db)

stringFileLoc <- "/mlab/data/rmflight/Documents/stringdatabase_91/"
```


# First question

* Of the genes that map to STRING, which ones map to AKT(??) and EGFR pathway? 
* any other pathways?

## Data

mRNA and miRNA from NSCLC cell lines with differential response to erolotinib (EGFR inhibitor). Affy U133 2.0, miRNA TaqMan card.

After reading in the data, what are our options?

* Use *STRINGdb* to get PPI (definite)
* Use *reactomedb* to get pathway membership (and consider co-occurence in pathway as an edge)
* Use *KEGGREST* and *KEGGGraph* to get pathway links between genes

stringdb - http://www.bioconductor.org/packages/release/bioc/vignettes/STRINGdb/inst/doc/STRINGdb.pdf
reactomedb

LOO cross-validated PLS results giving interactions between miRNA and mRNA.

```{r loadData}
mirData <- read.table("compiledinteractiondata.csv", header=TRUE, stringsAsFactors=FALSE, sep=",")
newmRNA <- sapply(seq(1, nrow(mirData)), function(x){
  useID <- mirData$mRNA[x]
  nID <- nchar(useID)
  substring(useID, 2, nID)
})
mirData$mRNA <- newmRNA
```

What is the distribution of associations of miRNA's to mRNA from the PLS results?

```{r linkDistribution}
splitMRNA <- split(mirData$miRNA, mirData$mRNA)
splitCount <- sapply(splitMRNA, length)
hist(splitCount)
```

Based on this (and comments from P Black), we will restrict our analysis to those Affy probesets with at least **5** miRNA associations.

```{r filterMRNA}
nMin <- 5
sum(splitCount >= nMin)

useMRNA <- data.frame(gene=names(splitCount)[splitCount >= nMin], stringsAsFactors=FALSE)
```

We want to convert our *Affy* ID's to other gene symbols and names for further use.

```{r convertID}
outMRNA <- select(hgu133plus2.db, useMRNA$gene, c("SYMBOL", "GENENAME", "ENTREZID"), "PROBEID")
```

## STRING PPI

We are doing a programmatic query of STRING because the website is rather limiting in how much data it will show you. So even though we will initially only look at direct interactions, we may get more information than is available off the website, as the website initially limits you to 10 interactions unless you explicitly ask it for more.

### Translate to STRING ID

Because we don't have direct access to STRING's translations, lets use a STRINGdb instance to do mapping of SYMBOLs to STRING IDs.

```{r setupSTRINGdb}
string_db <- STRINGdb$new(version="9_1", species=9606, input_directory=stringFileLoc)
outMRNA_mapped <- string_db$map(outMRNA, "SYMBOL", removeUnmappedRows=TRUE)
outMRNA_mapped$ENSEMBLPROT <- substring(outMRNA_mapped$STRING_id, 6, 20)
```

We've previously downloaded the STRINGdb files for human. Let's load them up and parse them down.

```{r stringData}
stringData <- read.table(file.path(stringFileLoc, "9606.protein.links.detailed.v9.1.txt"), header=TRUE, sep=" ", stringsAsFactors=FALSE)
head(stringData)
```

We will use the cutoff for `combined_score` of 400 (the default in STRING itself), and then filter down to just those things with co-expression, experimental or database evidence.

```{r filterData}
combScore <- 400
minScores <- c(experimental=400, database=400, coexpression=400)
passesScore <- rep(FALSE, nrow(stringData))
invisible(lapply(names(minScores), function(evidence){
  passesScore <<- passesScore | (stringData[, evidence] >= minScores[evidence])
}))
passesScore <- passesScore & (stringData$combined_score >= combScore)
head(stringData[passesScore,])
```

```{r trueFilter}
stringData <- stringData[passesScore,]
```

Now that we have possible interactions, we need to filter them down to links that involve our actual proteins of interest, and any links between those proteins.

```{r getUsefulLinks}
rnd1 <- which((stringData$protein1 %in% outMRNA_mapped$STRING_id) | (stringData$protein2 %in% outMRNA_mapped$STRING_id))

useNodes <- unique(c(stringData$protein1[rnd1], stringData$protein2[rnd1]))

hasP1 <- stringData$protein1 %in% useNodes
hasP2 <- stringData$protein2 %in% useNodes

useEdges <- hasP1 & hasP2

stringEdges <- stringData[useEdges,]
nrow(stringEdges)

# how many of our genes are actually in the network?
sum(outMRNA_mapped$STRING_id %in% unique(c(stringEdges$protein1, stringEdges$protein2)))
```

Now we should instantiate the network, and start adding information to it.

```{r createNetwork}
allNodes <- unique(c(stringEdges$protein1, stringEdges$protein2))
stringGraph <- graphNEL(nodes=allNodes, edgemode="directed")
stringGraph <- addEdge(stringEdges$protein1, stringEdges$protein2, stringGraph)
```

We also need to know some other stuff about the nodes.

```{r otherNodeData}
trimNodes <- substring(allNodes, 6, 20)
nodeInfo <- select(org.Hs.eg.db, trimNodes, c("SYMBOL", "GENENAME", "ENTREZID"), keytype="ENSEMBLPROT")
nodeInfo$nodeID <- paste("9606.", nodeInfo$ENSEMBLPROT, sep="")

# this last section is to make sure we are using the info from the affy chip for those nodes we already had.
hasNodes <- nodeInfo$nodeID %in% outMRNA_mapped$STRING_id
nodeInfo <- nodeInfo[!hasNodes,]

outMRNA_mapped$nodeID <- outMRNA_mapped$STRING_id
nodeInfo <- rbind(nodeInfo, outMRNA_mapped[,c("ENSEMBLPROT", "SYMBOL", "GENENAME", "ENTREZID", "nodeID")])
nodeInfo <- nodeInfo[(nodeInfo$nodeID %in% nodes(stringGraph)),]
```

We are also going to indicate which nodes originated from our original measurements

```{r addChipInfo}
nodeInfo$fromChip <- as.character(nodeInfo$nodeID %in% outMRNA_mapped$nodeID)
```

And add all that lovely data to our graph.

```{r addNodeData}
for (iName in c("ENSEMBLPROT", "SYMBOL", "GENENAME", "ENTREZID", "fromChip")){
  nodeDataDefaults(stringGraph, iName) <- ""
  attr(nodeDataDefaults(stringGraph, iName), "class") <- "STRING"
  nodeData(stringGraph, nodeInfo$nodeID, iName) <- nodeInfo[,iName]
}
```

Now lets see what we can see.

```{r loadCytoscape}
cw <- CytoscapeWindow("stringgraph", stringGraph)
displayGraph(cw)
layoutNetwork(cw, "force-directed")
```

Where are the original genes?

```{r selectOriginal}
orgNodes <- nodeInfo$nodeID[as.logical(nodeInfo$fromChip)]
selectNodes(cw, orgNodes, preserve.current.selection=FALSE)
```

OK, where do the AKT(1, 2, 3) genes show up?


