# Network Analysis

## Data

mRNA and miRNA from NSCLC cell lines with differential response to erolotinib (EGFR inhibitor). Affy U133 2.0, miRNA TaqMan card.

After reading in the data, what are our options?

* Use *STRINGdb* to get PPI (definite)
* Use *reactomedb* to get pathway membership (and consider co-occurence in pathway as an edge)
* Use *KEGGREST* and *KEGGGraph* to get pathway links between genes

stringdb - http://www.bioconductor.org/packages/release/bioc/vignettes/STRINGdb/inst/doc/STRINGdb.pdf
reactomedb

LOO cross-validated PLS results giving interactions between miRNA and mRNA.

```{r loadData}
mirData <- read.table("compiledinteractiondata.csv", header=TRUE, stringsAsFactors=FALSE, sep=",")
newmRNA <- sapply(seq(1, nrow(mirData)), function(x){
  useID <- mirData$mRNA[x]
  nID <- nchar(useID)
  substring(useID, 2, nID)
})
mirData$mRNA <- newmRNA
```

```{r linkDistribution}
splitMRNA <- split(mirData$miRNA, mirData$mRNA)
splitCount <- sapply(splitMRNA, length)
hist(splitCount)
sum(splitCount >= 5)

useMRNA <- names(splitCount)[splitCount >= 5]
```



```{r setupSTRINGdb}
library(STRINGdb)
string_db <- STRINGdb$new(version="9_1", species=9606)
```

